DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-02-01 paper/overview.tex
..

In this section, we provide an overview of our DeepPicar platform. In
developing DeepPicar, one of our primary goals is to faithfully
replicate NVIDIA's DAVE-2 system on a smaller scale with using a low cost
multicore platform, the Raspberry Pi 3. Because the Raspberry Pi 3's
computing performance is much lower than that of the DRIVE PX platform
used in DAVE-2, we are interested in if, and how, we can process
computationally expensive neural network operations in
real-time. Specifically, inferencing (forward pass processing)
operations must be completed within each control period
duration---e.g., a WCET of 33.$\overline{\mbox{33}}$ ms for 30Hz control 
frequency---locally on the Pi 3 platform, although training of the 
network (back-propagation for weight updates) can be done offline and 
remotely using a desktop computer.

@@

Figure~\ref{fig:overview} shows the DeepPicar, which is comprised of a
set of inexpensive components: a Raspberry Pi 3 Single Board Computer
(SBC), a Pololu DRV8835 motor driver, a Playstation Eye webcam, a
battery, and a 1:24 scale RC car. Table~\ref{tbl:carbom} shows the
estimated cost of the system.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.7\textwidth]{figs/architecture}
  \caption{DeepPicar's neural network architecture: 8 layers (5
    convolutional, 3 fully-connected layers), 27 million connections,
    250K parameters. The architecture is almost identical to the one 
	used in NVIDIA's real self-driving car~\cite{Bojarski2016}.}
  \label{fig:architecture}
\end{figure}

For the neural network architecture, we implement
NVIDIA DAVE-2's convolutional neural network (CNN) using TensorFlow,
based on ~\cite{deeptesla}. Note, however, that the CNN model
in~\cite{deeptesla} was considerably bigger than NVIDIA's CNN as it contained an
additional fully-connected layer of approximately 1.3M additional
parameters. We eliminated the additional layer to faithfully recreate
the NVIDIA's original CNN model.
%% The main difference is that we do not utilize a normalization layer, and 
%% instead initialize the weights using a Xavier initialization. 
As in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
as input and produces a single steering angle value as
output. Figure~\ref{fig:architecture} shows the network architecture
used in this paper, which
is comprised of 8 layers, 250K parameters, and about 27 million
connections. It is identical to NVIDIA DAVE-2's CNN except that we do
not apply a normalization layer as it does not include trainable
parameters and its computational demand is minimal.

\begin{figure}[h]
  \centering
  \includegraphics[width=.65\textwidth]{figs/track_new2}
  \caption{One of the custom tracks used for training/testing.}
  \label{fig:track}
\end{figure}

% data collection and training.
To collect the training data, a human pilot manually drives the RC car
on a small track we created (Figure~\ref{fig:track}) to record
timestamped videos and contol commands. The stored data is then copied 
to a desktop computer, which is equipped with a NVIDIA GTX 1060 GPU, 
where we train the network to accelerate training speed. 
For comparison, training the network on the Raspbeery Pi 3 takes
approximately 4 hours, whereas it takes only about 5 minutes on the
desktop computer using the GTX 1060 GPU.

@@

Another issue we observed in training/testing the network, which could
affect the network performance, is camera latency. In DeepPicar's
context, the camera latency is from the time the camera
sensor observes the scene to the time the computer actually reads the
digitized image data. This time can be noticable depending on the camera 
used and the performance of the Pi. We experimentally measured the camera
latency and found it to be around 50-100 ms. This is higher than the 
latency of human perception, which is known to be as fast as 13 
ms~\cite{ThomasBurger2015}. Higher camera latency could negatively 
affect control performance, because the DNN would analyze stale scenes. 
In the future, we plan to identify and use low-latency cameras.

% https://www.pubnub.com/blog/2015-02-09-how-fast-is-realtime-human-perception-and-technology/

Despite these issues, the trained models were still able to
achieve a reasonable degree of accuracy, successfully navigating
several different tracks we trained. In our testing, we found that the
DeepPicar could remain on a track for over 10
minutes~\footnote{A collection of self-driving videos can be found at: \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}}.

>>>>

@@

@@

In this section, we provide an overview of our DeepPicar platform.
In developing DeepPicar, one of our primary goals is to faithfully
replicate NVIDIA's DAVE-2 system on a smaller scale with using a low cost
multicore platform, the Raspberry Pi 3. Because the Raspberry Pi 3's
computing performance is much lower than that of the DRIVE PX platform
used in DAVE-2, we are interested in if, and how, we can process
computationally expensive neural network operations in
real-time. Specifically, inferencing (forward pass processing)
operations must be completed within each control period
duration---e.g., a WCET of 33.$\overline{\mbox{33}}$ ms for 30Hz control 
frequency---locally on the Pi 3 platform, although training of the 
network (back-propagation for weight updates) can be done offline and 
remotely using a desktop computer.
Figure~\ref{fig:overview} shows the DeepPicar, which is comprised of a
set of inexpensive components: a Raspberry Pi 3 Single Board Computer
(SBC), a Pololu DRV8835 motor driver, a Playstation Eye webcam, a
battery, and a 1:24 scale RC car. Table~\ref{tbl:carbom} shows the
estimated cost of the system.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.65\textwidth]{figs/architecture}
  \caption{DeepPicar's neural network architecture: 8 layers (5
    convolutional, 3 fully-connected layers), 27 million connections,
    250K parameters. The CNN architecture is identical to the one 
	used in NVIDIA's real self-driving car~\cite{Bojarski2016}.}
  \label{fig:architecture}
\end{figure}

For the neural network architecture, we implement
NVIDIA DAVE-2's convolutional neural network (CNN) using a TensorFlow
based CNN model in ~\cite{deeptesla}. Note, however, that the CNN model
in~\cite{deeptesla} was considerably bigger than NVIDIA's CNN model as
it contained an additional fully-connected layer of approximately 1.3M
additional parameters. We eliminated the additional layer to
faithfully recreate the NVIDIA's original CNN model.
%% The main difference is that we do not utilize a normalization layer, and 
%% instead initialize the weights using a Xavier initialization. 
As in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
as input and produces a single steering angle value as
output. Figure~\ref{fig:architecture} shows the network architecture
used in this paper, which is comprised of 8 layers, 250K parameters,
and about 27 million connections as in NVIDIA DAVE-2's. Note, however,
that we did not apply the normalization mentioned
in~\cite{Bojarski2016}, as it does not include trainable parameters
and its computational demand with respect to the overall CNN
processing is minimal.

\begin{figure}[h]
  \centering
  \includegraphics[width=.5\textwidth]{figs/track_new2}
  \caption{One of the custom tracks used for training/testing.}
  \label{fig:track}
\end{figure}

% data collection and training.
To collect the training data, a human pilot manually drives the RC car
on tracks we created (Figure~\ref{fig:track}) to record
timestamped videos and contol commands. The stored data is then copied 
to a desktop computer, which is equipped with a NVIDIA GTX 1050ti GPU, 
where we train the network to accelerate training speed. 
For comparison, training the network on the Raspbeery Pi 3 takes
approximately 4 hours, whereas it takes only about 4 minutes on the
desktop computer using the GTX 1050ti GPU.

@@

Other issues we observed in training/testing the network, which
can affect the prediction accuracy of the CNN, are camera and actuator
(motor) control latencies. In DeepPicar's
context, the camera latency is from the time the camera
sensor observes the scene to the time the computer actually reads the
digitized image data. This time can be noticable depending on the camera 
used and the performance of the Pi. Higher camera latency could
negatively affect control performance, because the DNN would analyze
stale scenes. Actuator (motor) control latency also can be a
factor, because moving the sterring motor to a desired position takes
time. 
In our platform, the combined latency is measured to be around
50 ms. In other words, in the recorded video and control data, a
control action (left, center, right) taken by the CNN appears to be
applied 50 ms later. Considering that camera's framerate is 30Hz
(33.3ms/frame), this is about two frames after the control action.
%% We experimentally measured the camera
%% latency and found it to be around 50-100 ms.
If this value is too high, control performance may suffer. 
For reference, the latency of human perception is known to be as fast
as 13 ms~\cite{ThomasBurger2015}.
% https://www.pubnub.com/blog/2015-02-09-how-fast-is-realtime-human-perception-and-technology/

Despite these issues, the trained models were still able to
achieve a reasonable degree of accuracy, successfully navigating
several different tracks we trained. In our testing, we found that the
DeepPicar could remain on a fairly challenging track for over 10
minutes at a moderate speed (50\% throttle), at which point we stop
the experiment, and more than one minute at a higher speed (75\%
throttle)~\footnote{A collection of self-driving videos can be found
  at: \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}}. Running at
higher speed is inherently more challenging because the CNN controller
has less time to recover from mistakes (bad predictions).
